Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-08-15 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
examining fds data & editting fire pattern algorithms

The fire pattern experiments were conducted in residential scale compartments built for the purpose of these experiments.  The compartments were constructed on top of a portable wood floor platform.  Each of the test compartments consisted of three 3.6~m long by 2.4~m high wood framed walls. The wood elements of the frame were composed of kiln-dried fir with a nominal cross section of 90~mm by 38~mm.  A partial ceiling with a width of at least 1.2~m, measured from the back wall of the compartment was installed across the width of the compartment, this allowed for the plume to impact the ceiling and form a ceiling jet along the back wall, before the fire gases flowed up and out of the compartment and into the 6~m by 6~m hood in the NIST Large Fire Research Laboratory. The interior surface of the walls and ceiling was composed of 12.7~mm thick, regular core gypsum wallboard. As defined by ASTM, gypsum wallboard is designed for use on walls, ceilings, or partitions.  It is composed of a non-combustible core, essentially gypsum, with paper bonded to the surface of the core. The back wall of the compartment consisted of three full gypsum board panels, each 1.2~m wide and 2.4~m high.  The front wall of the compartment was open along the full width of the compartment to minimize the impact of entrained air being channeled through a doorway. These experiments were conducted without instrumentation in the walls for several reasons:
\begin{enumerate}
\item to compare the non-instrumented gypsum wallboard fire pattern with the instrumented gypsum wallboard fire pattern results to determine if the instrumentation affected the pattern generation
\item to provide more realistic conditions
\item to enable the conduct of additional replicate experiments
\end{enumerate}

@@

In this chapter the application of emprical algebraic fire characterizations will be examined. Two different types of applications will be used.  In the first case relationships between the energy generated by a given fire source and the potential to transfer enough heat to generate thermal damage on a target, in this case gypsum board, will be examined.  In the second case, the height of the thermal damage on the gypsum board will be examined.  
Applications of algorithms such as these are presented in cureent fire investigaton textbooks~\cite{Icove:2013}.

@@

The first              ~\cite{Babrauskas:1981}

Ratio from heat flux from source fire data
vyto second item ignite

>>>>

The fire pattern experiments were conducted in residential scale compartments built for the purpose of these experiments.  The compartments were constructed on top of a portable wood floor platform.  Each of the test compartments consisted of three 3.6~m long by 2.4~m high wood framed walls. The wood elements of the frame were composed of kiln-dried fir with a nominal cross section of 90~mm by 38~mm.  A partial ceiling with a width of at least 1.2~m, measured from the back wall of the compartment was installed across the width of the compartment, this allowed for the plume to impact the ceiling and form a ceiling jet along the back wall, before the fire gases flowed up and out of the compartment and into the 6~m by 6~m hood in the NIST Large Fire Research Laboratory. The interior surface of the walls and ceiling was composed of 12.7~mm thick, regular core gypsum wallboard. As defined by ASTM, gypsum wallboard is designed for use on walls, ceilings, or partitions.  It is composed of a non-combustible core, essentially gypsum, with paper bonded to the surface of the core. The back wall of the compartment consisted of three full gypsum board panels, each 1.2~m wide and 2.4~m high.  The front wall of the compartment was open along the full width of the compartment to minimize the impact of entrained air being channeled through a doorway. These experiments were conducted without instrumentation in the walls for several reasons:
\begin{enumerate}
\item to compare the non-instrumented gypsum wallboard fire pattern with the instrumented gypsum wallboard fire pattern results to determine if the instrumentation affected the pattern generation
\item to provide more realistic conditions
\item to conduct additional replicate experiments
\end{enumerate}

@@

In this chapter the application of emprical algebraic fire characterizations will be examined. Two different types of applications will be used.  In the first case relationships between the energy generated by a given fire source and the potential to transfer enough heat to generate thermal damage on a target, in this case gypsum board, will be examined.  In the second case, the height of the thermal damage on the gypsum board will be examined.  
Applications of algorithms such as these are presented in current fire investigaton textbooks~\cite{Icove:2013}.

@@

The first algorithm assumes that the heat release rate is emitted from a point source in a uniform manner.  Therefore for any given radii a sphere of uniform heat flux will be assumed. A schematic of this arrangement is shown... and the equation is presented below. Babrauskas applied this concept in his paper, "Will the Second Item Ignite"~\cite{Babrauskas:1981}.  The idea being that if you knew the threshold heat flux that a material needed to be exposed to for the item to ignite, then you could potentially determine the maximum distance that the material could be located from a fire of a given HRR, with a given radiant fraction and it would still ignite.
